A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-08-07 graph_syntax.tex
files for submitted version and reviewer 3's comments

\section{Graph-Based Syntax Definition and Formalisms}
  \label{s:syntax definition}
  In this section we present our graph-based syntax, beginning with the necessary graph theoretic fundamentals to 
  construct the graphs discussed in Sec.~\ref{s:requirements}. 
  The notation used in this paper is adapted from Diestel \cite{Diestel2010}. 
  A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
  which means that the elements of $E$ are 2-element subsets of $V$. The set $V$ 
  contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
  For a \emph{directed graph} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
  of a set of sets. Each ordered pair represents an edge starting at the node 
  indicated by the first entry and directed to the node indicated by the second 
  entry. An edge $e$ = $(x,y)$ may be referred to simply as $xy$. The edges directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are given 
  by $E^{-1}(v)$. 
  If $E$ is not a one-to-one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
  \begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=1.5in]{images/example_directed_graph}
    \end{center}
    \vspace{-20pt}
  \caption{Example directed graph.}
  \label{f:example directed graph}
  \end{figure}
  As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
  \begin{IEEEeqnarray*}{rCl}
  V & = & \{1,2,3,4,5\}, \\
  E & = & \big\{(2,1),(3,2),(1,3),(3,4)\big\}.
  \end{IEEEeqnarray*}

@@

  %does this belong in the analysis block section???? 
  Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
  The set of sets $\mathcal A = \{A_i \st i \in I\}$ is called an indexed family of sets with index $i$ and 
  indexing set $I$\cite{smith2006}. 
  The union over this family of sets can be described in a few different ways:
  \begin{equation}
  \bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for some } A \in \mathcal A\}.
  \end{equation}

@@

\subsection{Node and Edge Types}
  \label{ss:types}
Along with these fundamentals of graph theory, we categorize nodes and edges into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
  \begin{description}
    \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
    \item[\bf{function node:}] represents the computation peformed by analysis tools,
    \item[\bf{driver node:}] control logic capable of managing iterations (PSG),
  \end{description}
and the three edge types are

  \begin{description}
  \item[\bf{connection edge:}] exchange of information external to analysis tools,
  \item[\bf{function edge:}] exchange of information internal to analysis tools,
  \item[\bf{driven edge:}] passing of information from a driver node to a variable node (PSG).
  \end{description}
Figure \ref{f:sellar types} demonstrates the usage of these node and edge types via the Sellar problem. function nodes are indicated by squares, variable nodes are indicated by circles, connection edges are indicated by dotted lines, and function edges are indicated by solid lines.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=4.0in]{images/sellar_types}
  \end{center}
  \caption{Sellar problem represented as a graph.}
\label{f:sellar types}
\end{figure}

  A rule set is provided for the usage of these node and edge types to provide a structure to the graph-based representation.
  The driver node and the driven edge are allowed to be present only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
  subject to the following restrictions: 
  \begin{enumerate}
  \item A function node may have only one edge directed to or from another function node.
  \item A function node may have only function edges directed in or out.
  \item A function node must have at least one edge directed in and at least one edge 
    directed out.
  \item If a variable node has an outgoing function edge, then it may not have other outgoing edges.
  \item If a variable node has an incoming function edge, then it may not have any additional incoming edges.
  \end{enumerate}

  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
  and one edge type are included \cite{alexandrov2004}. The present work adds one additional 
  node type and two additional edge types to the syntax to allow descriptive
  graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.

\subsection{Analysis Blocks}
\label{ss:analysis blocks}
Analysis tools take in a set of input variables and then calculate 
the values for their respective outputs. We represent this process
by a digraph called an \emph{analysis block}; a notional analysis block is shown in Fig.~\ref{f:analysis block}. 
%Each analysis block is a directed graph denoted as $ A=(V_{A},E_{A})$
As indicated in this figure, analysis blocks comprise three sets of nodes representing the distinct local inputs, the analysis (computation), and the distinct local outputs. The local input and local outputs nodes are variable nodes, while the nodes representing the analysis are function nodes. All of the edges within the analysis block are function edges, and are considered fixed to the analysis block. This graph structure demonstrated by Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}. Reciprocally, the rules ensure that analysis tools are represented via this analysis block structure.
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=3.0in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block with node types labeled.}
\label{f:analysis block}
\end{figure}

The function edge connecting the two function nodes represents the necessary calculations to map the inputs to the outputs of the analysis. 
This edge also provides the opportunity to encode computational cost or other characteristics of the analysis code as a weight in a weighted graph formulation. 
Since all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
The connectivity of nodes and edges in an analysis block cannot be altered for the purposes of MDAO problem formulation; 
however, analysis block sub-graphs can be added or removed from the MDAO problem graph as needed.
Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. 
As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a function node, and outputs are represented as variable nodes that have an incoming edge from a function node.

\subsection{Objectives, Constraints, and Inputs}
\label{ss:objectives and constraints}
Along with analysis tools, objectives, constraints, and inputs also need to be represented.
In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly multiple values are assembeled together to form a composite objective function. 
Generally, both objective and constraint functions accept a set of inputs and map them to an output value of significance to the overall design problem.

The operations of implementing composite objective and constraint functions, although typically simple, are effectively identical to the task performed by an analysis block. 
A composite constraint or objective function can therefore be represented as an analysis block within the graph with its own input and output variable nodes. 
Although fundamentally no different than an analysis block, for clarity and convience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph. 
Therefore, we use an \emph{expression block} to represent an objective or constraint. 
These expression block graphs follow the same structure as analysis blocks presented in Sec.~\ref{ss:analysis blocks}.

Finally, inputs to the problem formulation are represented in the graph by individual variable nodes with connection edges directed out to corresponding local inputs of analysis blocks, and they are referred to simply as \emph{inputs}. 
These nodes serve to indicate that the variable nodes they are connected to represent the same variable in the problem formulation.
These nodes serve to unify the variable nodes they are connected to. That is, they indicate that these nodes refer to the same input variable.

Figure \ref{f:obj-cons} demonstrates the use of inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=4.0in]{images/sellar_obj_and_cons}
  \end{center}
  \caption{Sellar problem graph indicating inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
\label{f:obj-cons}
\end{figure}

\subsection{Indegree and Outdegree Limits}
  \label{s:indegree-outdegree}
	To address the MDAO notion of design variables, we first introduce the concept of the degree of a node.
  In a digraph, \emph{indegree} of a node is the number of edges directed in and 
  is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
  is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$ \cite{Diestel2010}.
  In this paper, the indegree and outdegree of a given node refer only to the number of the connection edges 
  directed into or out of the node, respectively.

  We may now define the \emph{upper indegree limit} 
  \begin{equation}
  \txt{deg}_u^-(v):V \to \mathbb{N}
  \end{equation} 
  and the \emph{lower indegree limit} as
  \begin{equation}
  \txt{deg}_l^-(v):V \to \mathbb{N}.
  \end{equation}
  These user-specified limits govern the number of connection edges that may 
  be directed into a node for a valid graph; of course, $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$ must be satisfied. 
For example, consider a variable 
  node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
  must have exactly one incoming connection edge or the graph is deemed an invalid problem formulation.

\section{MDAO Constructs Derived from the Graph-Based Syntax}
\label{s:graph representation}
In Sec.~\ref{s:syntax definition} we presented the structure of the graph-based representation. We now discuss how these formalisms provide the remaining MDAO problem constructs identified in Sec.~\ref{s:requirements}.

%To begin, we make use of the Sellar problem, described in 
%Eq. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
%The graph representation of the Sellar problem, using the new stynax, is 
%presented in Fig. \ref{f:sellar_graph_full}. In this section sub-graphs or 
%slight modifications of the full graph will be used to demonstrate the 
%important MDAO constructs described by this graph syntax.
%\begin{figure}[htb!]
%    \begin{center}
%    \includegraphics[width=4in]{images/sellar_graph_full}
%    \end{center}
%    \vspace{-10pt}
%\caption{Graph of the Sellar problem formulation}
%\label{f:sellar_graph_full}
%\end{figure}

\subsection{Design Variables, Holes, and Collisions}
Design variables are input variables whose values are free to be 
changed by the designer or by an optimizer in an MDAO study. In the graph-based syntax, a node corresponds to a design variable when the indegree limit of the node is prescribed as zero.

If the indegree limits are violated by the number of incoming connection edges (the indegree of the node), the graph is regarded as an invalid problem formulation. There are two classifications for these violations:
  \begin{description}
    \item[\bf{hole:}] The number of incoming edges is less than the lower indegree limit:
      \begin{equation} 
\txt{deg}^-(v) < \txt{deg}_l^-(v). 
\label{e:hole} 
\end{equation}
      A hole represents a lack of information being supplied to the variable node. This implies that the analysis tool being represented by the analysis block would not be capable of producing outputs.
    \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit. 
      \begin{equation} 
\txt{deg}^-(v) > \txt{deg}_u^-(v). 
\label{e:collision}
\end{equation}A collision represents redundant information being supplied to the variable node from two or more sources. This conflict implies an ambiguity as to which information is to be used as an input for the analysis tool.
  \end{description}

  The presence of holes and collisions in a graph represent conflicts that will give
  rise to an invalid problem formulation. It is expected that a maximal connectivity graph could indicate these conflicts, whereas a fundamental problem graph will not.

\subsection{Coupling Between Analyses}
  In MDAO, coupling is the mutual dependence the outputs of two or more analysis 
  codes. In this graph-based syntax, coupling is described by the presence of a 
  cycle between two or more analysis blocks. In the Sellar problem from Eqn \ref{eqn:simple_fpf} 
  a reciprical dependence between $D_1$ and $D_2$ exists through the variabled $y_1$ and $y_2$. 
  The edges belonging to a cycle in the Sellar problem are highlighted in 
  Fig. \ref{f:sellar cycles} where they form a closed path between analysis blocks $D_1$ and $D_2$.

 \begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=4.0in]{images/sellar_cycles}
  \end{center}
    \caption{Sellar problem graph with dashed edges indicating participation in a cycle.
\label{f:sellar cycles}
    }
\end{figure}

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be necessary to build a valid PSG from a given FPG. 
  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $D_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}.
  Larger problems can contain more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentailly reduce the number of cycles in a graph.

  If many couplings are present, convergence rates can be improved by 
  searching for an effective ordering for the execution of analyses.
  Rogers' DEMAID tool uses a genetic algorithm to find an ordering that minimizes 
  the overall coupling of the system by separating independent cycles in the 
  graph \cite{rogers1996,rogers1996demaid}. Rogers work focused on the matrix 
  form of the DSM for ordering optimization. Lu and Martins more recently leveraged 
  a weighted form of the DSM and used an iterative clustering approach to perform a 
  similar task to DEMAID \cite{Lu2012}.

\subsection{Multi-fidelity Problems}
  \label{ss:multi-fideliy problems}
  A multi-fidelity MDAO problem can be characterized by a formulation in which 
  two or more different analyses each calculating the same data. Multi-fidelity 
  is represented in a graph by a variable node having an indegree greater than 
  one, which means that multiple connection edges are directed into it. When the 
  upper indegree limit of a variable node is set above one, then the node is 
  implied to allow multiple fidelity instances of the variable. When the lower 
  indegree limit of a variable node is set above one, then multiple fidelity instances are required.

  In a multi-fidelity problem, a given variable node may have multiple incoming 
  edges without implying a collision as defined by by Eq.~\ref{e:collision}. 
An  upper indegree limit above unity for any variable in a graph represents a decision to 
  allow multiple fidelities to interact at that part of the graph in order to allieviate
  a conflict. 
  Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing these upper degree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
In section \ref{ss:obtaining FPG} we present an algorithm for finding 
  conflicts within a given MCG so that a designer can make the necessary decisisons about each 
  one in turn. 
  \begin{figure}
    \begin{center}
      \includegraphics[width=4in]{images/sellar_mulfi}
    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
  \end{center}
  \end{figure}

  Multi-fidelity problems are always charaterized by graphs with variable 
  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
  special techniques for resolving the confliciting edges by introducing some mechanism
  to manage when each of the different fidelity analyses are 
  run\cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
  The specifics of this mechanism are not given in any form withing an MCG or 
  FPG. Instead the multi-fidelity mechanism specifics would be represented as a 
  driver in a PSG derived from a multi-fidelity FPG.

Table \ref{t:variable node classification} provides examples to summarize the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity.

>>>>

\section{Graph-Based Syntax Definition and Formalisms}
  \label{s:syntax definition}
  In this section we present our graph-based syntax, beginning with the necessary graph theoretic fundamentals to 
  construct the graphs discussed in Sec.~\ref{s:requirements}. 
  The notation used in this paper is adapted from Diestel \cite{Diestel2010}. 
  A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
  which means that the elements of $E$ are 2-element subsets of $V$. The set $V$ 
  contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
  For a \emph{directed graph} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
  of a set of sets. Each ordered pair represents an edge starting at the node 
  indicated by the first entry and directed to the node indicated by the second 
  entry. An edge $e$ = $(x,y)$ may simply be referred to as $xy$. The edges 
  directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are denoted 
  by $E^{-1}(v)$. $E(v)$ may be the empty set, a single edge, or a set of edges, and likewise for $E^{-1}(v)$.
  \begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=1.5in]{images/example_directed_graph}
    \end{center}
    \vspace{-20pt}
  \caption{Example directed graph.}
  \label{f:example directed graph}
  \end{figure}
  As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
  \begin{IEEEeqnarray*}{rCl}
  V & = & \{1,2,3,4,5\}, \\
  E & = & \big\{(2,1),(3,2),(1,3),(3,4)\big\}.
  \end{IEEEeqnarray*}

@@

  %does this belong in the analysis block section???? 
  Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
  The set of sets $\mathcal A = \{A_i \st i \in I\}$ is called an indexed family of sets with index $i$ and 
  indexing set $I$ \cite{smith2006}. 
  The union over this family of sets can be described in a few different ways:
  \begin{equation}
  \bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for some } A \in \mathcal A\}.
  \end{equation}

@@

\subsection{Node and Edge Types}
  \label{ss:types}
We categorize nodes and edges 
  into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
    \begin{description}
      \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
      \item[\bf{function node:}] represents the computation peformed by analysis tools,
      \item[\bf{driver node:}] represents control logic capable of managing iterations (present only in a PSG),
    \end{description}
  and the three edge types are

  \begin{description}
  \item[\bf{connection edge:}] represents exchange of information external to analysis tools,
  \item[\bf{function edge:}] represents exchange of information internal to analysis tools,
  \item[\bf{driven edge:}] represents passing of information from a driver node to a variable node (present only in a PSG).
  \end{description}
  Figure \ref{f:sellar types} demonstrates the usage of these node and edge 
  types via the Sellar problem. 
In this figure, function nodes are indicated by squares, 
  variable nodes are indicated by circles, connection edges are indicated by 
  dotted lines, and function edges are indicated by solid lines.
  \begin{figure}[htb!]
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_types}
    \end{center}
    \caption{Sellar problem represented as a graph.}
  \label{f:sellar types}
  \end{figure}

  A rule set is provided for the usage of these node and edge types to provide a structure to the graph-based representation.
  The driver node and the driven edge are allowed only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
  subject to the following restrictions: 
  \begin{enumerate}
  \item A function node may have only one edge directed to or from another function node.
  \item A function node may have only function edges directed in or out.
  \item A function node must have at least one edge directed in and at least one edge 
    directed out.
  \item If a variable node has an outgoing function edge, then it may not have other outgoing edges.
  \item If a variable node has an incoming function edge, then it may not have any additional incoming edges.
  \end{enumerate}

 Alexandrov and Lewis's REMS syntax includes only two node types (variable and function) 
  and one edge type   \cite{alexandrov2004}. The present work adds one additional 
  node type and two additional edge types to the syntax to allow descriptive
  graphs for all three phases of the design problem formulation process discussed in Sec. \ref{s:requirements}.

\subsection{Analysis Blocks}
\label{ss:analysis blocks}
  Analysis tools take in a set of input variables and calculate 
  the values for their respective outputs. We represent this process
  by a digraph called an \emph{analysis block}; a notional analysis block is shown in Fig.~\ref{f:analysis block}. 
  %Each analysis block is a directed graph denoted as $ A=(V_{A},E_{A})$
  As indicated in this figure, analysis blocks comprise three sets of nodes 
  representing the distinct local inputs, the analysis (computation), and the 
  distinct local outputs. The local input and local output nodes are variable 
  nodes, while the nodes representing the analysis are function nodes. All of 
  the edges within the analysis block are function edges, and are considered 
  fixed to the analysis block. This graph structure demonstrated by 
  Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}. 
  Conversely, the rules ensure that analysis tools are represented via this analysis block structure.

  \begin{figure}[htb]
      \begin{center}
      \includegraphics[width=3.0in]{images/analysis_block}
      \end{center}
      \vspace{-10pt}
  \caption{Example analysis block with node types labeled.}
  \label{f:analysis block}
  \end{figure}

  The function edge connecting the two function nodes represents the necessary 
  calculations to map the inputs to the outputs of the analysis. 
  This edge also provides the opportunity to encode computational cost or other 
  characteristics of the analysis code as a weight in a weighted graph formulation. 
  Because all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
  The connectivity of nodes and edges in an analysis block cannot be altered for the purposes of MDAO problem formulation; 
  however, analysis block sub-graphs can be added or removed from the MDAO problem graph as needed.
  Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. 
  As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a function node, and outputs are represented as variable nodes that have an incoming edge from a function node.

\subsection{Objectives, Constraints, and Inputs}
  \label{ss:objectives and constraints}
  Along with analysis tools, objectives, constraints, and inputs also need to be represented.
  In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly, multiple output values from different analysis tools are aggregated together to form a composite objective function. 
  Generally, both objective and constraint functions accept a set of inputs and map them to an output value of significance to the overall design problem.

  The operations of implementing composite objective and constraint functions, although typically simple, are effectively identical to the task performed by an analysis block. 
  A composite constraint or objective function can therefore be represented as an analysis block within the graph with its own input and output variable nodes. 
  Although fundamentally no different than an analysis block, for clarity and convenience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph. 
  Therefore, we use an \emph{expression block} to represent an objective or constraint. 
  These expression block graphs follow the same structure as analysis blocks presented in Sec.~\ref{ss:analysis blocks}.

  Finally, inputs to the problem formulation are represented in the graph by individual variable nodes with connection edges directed out to corresponding local inputs of analysis blocks, and they are referred to simply as \emph{inputs}. 
  These nodes serve to indicate that the variable nodes within analysis or expression blocks to which they are connected represent the same variable in the problem formulation.

  Figure \ref{f:obj-cons} demonstrates the use of inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
  \begin{figure}[htb!]
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_obj_and_cons}
    \end{center}
    \caption{Sellar problem graph indicating inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
  \label{f:obj-cons}
  \end{figure}

\subsection{Indegree and Outdegree Limits}
  \label{s:indegree-outdegree}
	To address the MDAO notion of design variables, we first introduce the concept of the degree of a node.
  In a digraph, \emph{indegree} of a node is the number of edges directed in and 
  is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
  is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$ \cite{Diestel2010}.

  We may now define the \emph{upper indegree limit} 
  \begin{equation}
  \txt{deg}_u^-(v):V \to \mathbb{N}
  \end{equation} 
  and the \emph{lower indegree limit} as
  \begin{equation}
  \txt{deg}_l^-(v):V \to \mathbb{N}.
  \end{equation}
  These user-specified limits govern the number of connection edges that may 
  be directed into a node for a valid graph; of course, the specification must satisfy $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$.
  For example, consider a variable node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
  must have exactly one incoming connection edge or the graph is deemed an invalid problem formulation.

\section{MDAO Constructs Derived from the Graph-Based Syntax}
  \label{s:graph representation}
  In Sec.~\ref{s:syntax definition}, we presented the structure of the graph-based representation. We now discuss how these formalisms provide the remaining MDAO problem constructs identified in Sec.~\ref{s:requirements}.

\subsection{Design Variables, Holes, and Collisions}
  Design variables are input variables whose values are free to be 
  changed by the designer or by an optimizer in an MDAO study. In the graph-based 
  syntax, a node corresponds to a design variable if the indegree limits of the node are prescribed as zero.

  If the indegree limits are violated by the number of incoming connection edges 
  (the indegree of the node), then the graph is regarded as an invalid problem formulation. There are two classifications for these violations:
    \begin{description}
      \item[\bf{hole:}] The number of incoming edges is less than the lower indegree limit:
        \begin{equation} 
      \txt{deg}^-(v) < \txt{deg}_l^-(v). 
      \label{e:hole} 
      \end{equation}
            A hole represents a lack of information being supplied to the variable node. 
            This implies that the analysis tool being represented by the analysis block 
            would not be capable of producing outputs.
          \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit:
            \begin{equation} 
      \txt{deg}^-(v) > \txt{deg}_u^-(v). 
      \label{e:collision}
      \end{equation}A collision represents redundant information being supplied to 
      the variable node from two or more sources. This conflict implies an 
      ambiguity as to which information is to be used as an input for the analysis tool.
	In multifidelity problems (discussed subsequently), the upper indegree limit could be specified as $deg_u^- > 2$, which implies that a collision is not noted in the case of multiple inputs.
  \end{description}

  The presence of holes and collisions in a graph represent conflicts that will give
  rise to an invalid problem formulation. 
The process of obtaining an FPG from an MCG will reveal these conflicts.

\subsection{Coupling Between Analyses}
  In MDAO, coupling is the mutual dependence between two or more analysis 
  tools and their respective outputs. In this graph-based syntax, coupling is described by the presence of a 
  cycle between two or more analysis blocks. In the Sellar problem from Eq.~\ref{eqn:simple_fpf} 
  a mutual dependence between $D_1$ and $D_2$ exists through the variables $y_1$ and $y_2$. 
  The edges belonging to a cycle in the Sellar problem are highlighted in 
  Fig. \ref{f:sellar cycles}.  These edges collectively form a closed path between analysis blocks $D_1$ and $D_2$.

 \begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=3.25in]{images/sellar_cycles}
  \end{center}
    \caption{Sellar problem graph with dashed edges indicating participation in a cycle.
\label{f:sellar cycles}
    }
\end{figure}

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be required to build a valid PSG from a given FPG  that includes cycles. 
  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select a
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $D_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}.
  Larger problems can involve more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentailly reduce the number of cycles in a graph.

  If many couplings are present, convergence rates can be improved by 
  searching for an effective ordering for the execution of analyses.
  Rogers' DeMAID tool uses a genetic algorithm to find an ordering that minimizes 
  the overall coupling of the system by separating independent cycles in the 
  graph \cite{rogers1996,rogers1996demaid}. Rogers work focused on the matrix 
  form of the DSM for ordering optimization. Lu and Martins more recently leveraged 
  a weighted form of the DSM and used an iterative clustering approach to perform a 
  similar task to DeMAID \cite{Lu2012}.

\subsection{Multi-fidelity Problems}
  \label{ss:multi-fideliy problems}
  A multi-fidelity MDAO problem can be characterized by a formulation in which 
  two or more different analyses each calculate the same data. Multi-fidelity 
  is represented in a graph by a variable node having an indegree greater than 
  one, which means that multiple connection edges are directed into it. When the 
  upper indegree limit of a variable node is set above one, then the node is 
  implied to allow multiple fidelity instances of the variable. When the lower 
  indegree limit of a variable node is set above one, then multiple fidelity instances are required.

  In a multi-fidelity problem, a given variable node may have multiple incoming 
  edges without implying a collision as defined by Eq.~\ref{e:collision}. 
  Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing the upper indegree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
In Sec.~\ref{ss:obtaining FPG} we present an algorithm for locating 
  conflicts within a given MCG such that a designer can make the necessary decisisons about each 
  one in turn. 
  \begin{figure}
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_mulfi}
    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
  \end{center}
  \end{figure}

  Multi-fidelity problems are always charaterized by graphs in which at least some of the variable 
  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
  special techniques for resolving the confliciting edges by introducing some mechanism
  to manage when each of the different fidelity analyses are 
  run \cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
  The specifics of this mechanism are not represented within an MCG or 
  FPG. Instead, the multi-fidelity mechanism specifics would be represented as a 
  driver in a PSG derived from a multi-fidelity FPG.

Table \ref{t:variable node classification} summarizes the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity node.
